Supporting Information for\\ ``Local fitness and epistatic effects lead to distinct patterns of linkage disequilibrium in protein-coding genes''

### 2022-06-03 manuscript/supplement.tex
updates for minor revisions

To explore the effects of many linked selected loci on patterns of LD, I used
\texttt{fwdpy11} \citep{Thornton2014-pn,Thornton2019-qc} to simulate large
genomic segments with many selected mutations. Each simulation had population
size of \(N=1000\), a sequence length of 1Mb, with mutation and recombination
rates both \(u=r=10^{-8}\). I considered four scenarios for selected mutaions:
additive selection \(h=0.5\) with \(s=-0.0005\) and \(s=-0.01\) (\(\gamma=-1\)
and \(-20\)), and recessivity \(h=0.0\) or \(h=0.1\) with the same selection
coefficients. These sets of parameters should give rise to strong effects of
linked selection, as there are many cosegregating, tightly linked selected
variants. For each setting, I examined site-frequency spectra and LD decay for
both selected and neutral mutations across the region.

@@

Mutations at different frequencies are expected to have different signs and
magnitudes of LD. For example, rare mutations show highly positive LD, even for
pairs of neutral mutations \citep{Good2022-ot}. Mutations of different
functional classes (synonymous, missense, and loss-of-function, for example)
are expected to segregate at different average frequencies, and average
selection pressures are observed to be greater within annotated conserved
domains, with stronger purifying selection driving allele frqeuencies to
relatively lower levels among missense and loss-of-function variants. Since
mutations of difference classes and locations have different average
frequencies, this may drive differences in patterns of LD when comparing
between them.

@@

\begin{longtable}[t]{lllr}
\caption{\label{tab:tajimasD}
\textbf{Tamija's $D$ for classes of coding mutations.} Here, we partitioned by
synonymous, missense, and nonsense mutationsi, and considered all mutations gene-wide,
mutations falling within annotated conserved elements,
and mutations falling outside of those
elements.}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endfirsthead
\caption[]{\label{tab:tajimasD}Tamija's $D$ for classes of coding mutations.
\textit{(continued)}}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endhead

@@

\CatchFileDef{\afrsigned}{AFR.sd1.tex}{}
\begin{table}
    \begin{center}
        \begin{tabular}{lccc}
            \afrsigned
        \end{tabular}
    \end{center}
    \caption{
        \label{tab:afrsd1}
        \textbf{Signed LD in AFR-labeled populations matched to distances
        and allele frequencies of synonymous mutations.}
        Gene-wide signed LD is observed to be at similar slightly positive levels between
        pairs of synonymous and pairs of missense variants. This observation is
        consistest when weighting missense variants to match distances between
        pairs of synonymous variants and when weighting to match synonymous allele
        frequencies, with only a slight reduction in missense \(\sigma_d^1\).
        The distribution of distances between synonymous and missense mutations
        within genes are very similar (Figure~\ref{fig:mutDistances}),
        so matching by distances separating mutations
        has a minor effect on measured average LD.
        The large measurement uncertainty of loss-of-function mutation pairs
        due to few observed pairs of such mutations within the same gene causes
        signed LD to fluctuate when conditioning on distances and frequencies
        matched to synonymous mutations.
    }
\end{table}

@@

\CatchFileDef{\eursigned}{EUR.sd1.tex}{}
\begin{table}
    \begin{center}
        \begin{tabular}{lccc}
            \eursigned
        \end{tabular}
    \end{center}
    \caption{
        \label{tab:eursd1}
        \textbf{Signed LD in EUR-labeled populations matched to distances
        and allele frequencies of synonymous mutations.}
        Similar to the pattern observed in AFR-labeled populations
        (Table~\ref{tab:afrsd1}), \(sigma_d^1\) between missense mutations matched
        to distances and allele frequencies observed in synonymous mutation pairs
        has only a small effect on average signed LD, and gene-wide measurements
        are roughy equal between these two mutation classes.
        Again, measurement noise for loss-of-function mutation pairs is high.
    }
\end{table}

@@

\CatchFileDef{\eassigned}{EAS.sd1.tex}{}
\begin{table}
    \begin{center}
        \begin{tabular}{lccc}
            \eassigned
        \end{tabular}
    \end{center}
    \caption{
        \label{tab:eassd1}
        \textbf{Signed LD in EUR-labeled populations matched to distances
        and allele frequencies of synonymous mutations.}
        Similar to the pattern observed in AFR-labeled populations
        (Table~\ref{tab:afrsd1}), \(sigma_d^1\) between missense mutations matched
        to distances and allele frequencies observed in synonymous mutation pairs
        has only a small effect on average signed LD, and gene-wide measurements
        are roughly equal between these two mutation classes. 
        Again, measurement noise for loss-of-function mutation pairs is high.
    }
\end{table}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_5e-06_s_0.0002_h_0.1.pdf}
    \caption{
        \textbf{Comparison to simulations with site-wise dominance.}
        \(Ne=5000\), \(r=5\times10^{-6}\), and \(s=-0.0002\) (\(\rho=0.1\) and
        \(\gamma=-2\)), with \(h=0.1\) at both loci and no
        epistatic interactions (\(\epsilon=0\)).
        E: \texttt{moments} expectations, D: simulated data.
        Additional comparisons to simulated data are availabe at
        \texttt{https://github.com/apragsdale/two\_locus\_selection}.
    }
    \label{fig:validation5}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with slightly deleterious mutations.}
        Using an individual based forward simulator \citep{Thornton2019-qc}, I
        simulated many replicates of 1Mb regions with high mutation rates of
        selected mutations (see Section~\ref{sec:bgs} for details). Here,
        many linked additive weakly selected mutations slightly distort the
        SFS for both selected and neutral mutations. \(\sigma_d^2\) is reduced
        relative to two-locus expectations within background selection, and
        Hill-Robertson interference effects are slightly diminished with
        \(\sigma_d^1\) not as strongly negative as expected.
    }
    \label{fig:bgs1}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with moderately deleterious mutations.}
        In the case of many linked strongly selected additive mutations, neither
        selected nor neutral statistics deviate dramatically from two-locus expectations.
        The SFS from the simulated data matches expectations, with a slight excess
        of selected mutations. LD is also largely unaffected, with \(\sigma_d^2\)
        matching expectations and \(\sigma_d^1\) fluctuating near zero as expected.
    }
    \label{fig:bgs2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.0_n_50}
    \caption{
        \textbf{Associative overdominance due to slightly deleterious recessive mutations.}
        The presence of many linked weakly selected recessive mutations can cause
        measues of diversity to have large increases over single- and two-locus
        expectations. The SFS for both neutral and selected mutations both show
        a very large excess of common variants (note the log scale), and \(\sigma_d^2\)
        are also increased over two-locus expectations for both classes of mutations.
        Still, \(\sigma_d^1\) remains 0 for neutral mutations and is not as strongly
        negative for pairs of selected mutations.
    }
    \label{fig:bgs3}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.1_n_80}
    \caption{
        \textbf{Background selection with moderately deleterious recessive mutations.}
        More strongly deleterious recessive mutations do not show as dramatic of
        deviations from expectations as the case of weakly deleterious recessive
        mutations. The neutral SFS is reduced, and we observe more selected mutations
        at higher frequencies than expected. LD is also only weakly distorted from
        two-locus expectations. Notably, we see positive LD (matching between
        simulations and expectations) for strongly deleterious recessive mutations,
        as predicted in \citet{Roze2021-cf}.
    }
    \label{fig:bgs4}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.sd2}
    \caption{
        \textbf{Time series of squared LD under bottleneck 
        and expansion histories with epistasis.}
        Bottlenecks increase \(\sigma_d^2\) (squared LD),
        while expansions tend to decrease squared LD.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance}
    \caption{
        \textbf{Time series of signed LD under bottleneck 
        and expansion histories with dominance.}
        Gene-based dominance, unlike site-wise dominance effects, can cause
        large positive signed LD, similar to compensatory mutations or
        antagonistic epistasis. In general, common variants are more
        stable under population size changes than rare variants.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_dom}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance.sd2}
    \caption{
        \textbf{Time series of squared LD under bottleneck 
        and expansion histories with dominance.}
        Similar to the cases with epistasis, bottlenecks tend to increase
        \(\sigma_d^2\) and expansions decrease squared LD.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_dom_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU}
    \caption{
        \textbf{Time series of signed LD under inferred models of human population
        size history with with epistasis.}
        Similar to the instantaneous size changes in the toy models
        (Figure~\ref{fig:toy}), bottlenecks tend to make signed LD more extreme.
        However, in the case of strong selection and antagonistic epistasis,
        the CEU \(\sigma_d^1\) trajector decreases relative to the YRI
        trajectory. Again, statistics for rare variants fluctuate more rapidly
        than common variants.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.sd2}
    \caption{
        \textbf{Time series of squared LD under inferred models of human population
        size history with with epistasis.}
        Size reductions, such as bottlenecks in the recent past of Eurasian populations,
        increase squared LD.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.dominance.sd2}
    \caption{
        \textbf{Time series of squared LD under inferred models of human population
        size history with with dominance.}
        Again, bottlenecks increase squared LD relative to populations that did
        not experience such strong size reductions in their history.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate_dom_sd2}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_in_domain}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        that fall inside the same domain.}
        At short distances between SNPs that both fall within sthe same conserved
        annotated domain, missense mtuations show consistently
        larger signed LD than synonymous variants. This may be caused by
        antagonistic epistasis or compensatory effects between pairs of
        nearby missense mutations.
    }
    \label{fig:LDwithin}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_afr}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes African group.
        We observe increased LD between missense mutations within domains,
        but decrased LD outside of domains at matched distances. This pattern
        holds when conditioning on non-rare variation (\(n_A, n_B \geq 9\),
        MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsAFR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eur}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes European group.
        We again observe increased LD between missense mutations within domains,
        but decrased LD outside of domains at matched distances. And again,
        this pattern holds when conditioning on non-rare variation
        (\(n_A, n_B \geq 9\), MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsEUR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eas}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes East Asian group.
        As with the African and European data,
        wobserve increased LD between missense mutations within domains,
        but decrased LD outside of domains at matched distances. And again,
        this pattern holds when conditioning on non-rare variation
        (\(n_A, n_B \geq 9\), MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsEAS}
\end{figure}

@@

>>>>

To explore the effects of many linked selected loci on patterns of LD, I used
\texttt{fwdpy11} \citep{Thornton2014-pn,Thornton2019-qc} to simulate large
genomic segments with many selected mutations. Each simulation had population
size of \(N=1000\), a sequence length of 1Mb, with mutation and recombination
rates both \(u=r=10^{-8}\). I considered four scenarios for selected mutations:
additive selection \(h=0.5\) with \(s=-0.0005\) and \(s=-0.01\) (\(\gamma=-1\)
and \(-20\)), and recessivity \(h=0.0\) or \(h=0.1\) with the same selection
coefficients. These sets of parameters should give rise to strong effects of
linked selection, as there are many cosegregating, tightly linked selected
variants. For each setting, I examined site-frequency spectra and LD decay for
both selected and neutral mutations across the region.

@@

Mutations at different frequencies are expected to have different signs and
magnitudes of LD. For example, rare mutations show highly positive LD, even for
pairs of neutral mutations \citep{Good2022-ot}. Mutations of different
functional classes (synonymous, missense, and loss-of-function, for example)
are expected to segregate at different average frequencies, and average
selection pressures are observed to be greater within annotated conserved
domains, with stronger purifying selection driving allele frequencies to
relatively lower levels among missense and loss-of-function variants. Since
mutations of difference classes and locations have different average
frequencies, this may drive differences in patterns of LD when comparing
between them.

@@

\begin{longtable}[t]{lllr}
\caption{\label{tab:tajimasD}
\textbf{Tamija's $D$ for classes of coding mutations.} Here, we partitioned by
synonymous, missense, and nonsense mutations, and considered all mutations gene-wide,
mutations falling within annotated conserved elements,
and mutations falling outside of those
elements.}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endfirsthead
\caption[]{\label{tab:tajimasD}Tamija's $D$ for classes of coding mutations.
\textit{(continued)}}\\
\toprule
Population & Mutation type & Region & Tajima's D\\
\midrule
\endhead

@@

\CatchFileDef{\afrsigned}{AFR.sd1.tex}{}
\begin{table}
    \begin{center}
        \begin{tabular}{lccc}
            \afrsigned
        \end{tabular}
    \end{center}
    \caption{
        \label{tab:afrsd1}
        \textbf{Signed LD in AFR-labeled populations matched to distances
        and allele frequencies of synonymous mutations.}
        Gene-wide signed LD is observed to be at similar slightly positive levels between
        pairs of synonymous and pairs of missense variants. This observation is
        consistent when weighting missense variants to match distances between
        pairs of synonymous variants and when weighting to match synonymous allele
        frequencies, with only a slight reduction in missense \(\sigma_d^1\).
        The distribution of distances between synonymous and missense mutations
        within genes are very similar (Figure~\ref{fig:mutDistances}),
        so matching by distances separating mutations
        has a minor effect on measured average LD.
        The large measurement uncertainty of loss-of-function mutation pairs
        due to few observed pairs of such mutations within the same gene causes
        signed LD to fluctuate when conditioning on distances and frequencies
        matched to synonymous mutations.
    }
\end{table}

@@

\CatchFileDef{\eursigned}{EUR.sd1.tex}{}
\begin{table}
    \begin{center}
        \begin{tabular}{lccc}
            \eursigned
        \end{tabular}
    \end{center}
    \caption{
        \label{tab:eursd1}
        \textbf{Signed LD in EUR-labeled populations matched to distances
        and allele frequencies of synonymous mutations.}
        Similar to the pattern observed in AFR-labeled populations
        (Table~\ref{tab:afrsd1}), \(sigma_d^1\) between missense mutations matched
        to distances and allele frequencies observed in synonymous mutation pairs
        has only a small effect on average signed LD, and gene-wide measurements
        are roughly equal between these two mutation classes.
        Again, measurement noise for loss-of-function mutation pairs is high.
    }
\end{table}

@@

\CatchFileDef{\eassigned}{EAS.sd1.tex}{}
\begin{table}
    \begin{center}
        \begin{tabular}{lccc}
            \eassigned
        \end{tabular}
    \end{center}
    \caption{
        \label{tab:eassd1}
        \textbf{Signed LD in EAS-labeled populations matched to distances
        and allele frequencies of synonymous mutations.}
        Similar to the pattern observed in AFR-labeled populations
        (Table~\ref{tab:afrsd1}), \(sigma_d^1\) between missense mutations matched
        to distances and allele frequencies observed in synonymous mutation pairs
        has only a small effect on average signed LD, and gene-wide measurements
        are roughly equal between these two mutation classes. 
        Again, measurement noise for loss-of-function mutation pairs is high.
    }
\end{table}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../simulations/discrete/plots/comp_Ne_5000_n_50_r_5e-06_s_0.0002_h_0.1.pdf}
    \caption{
        \textbf{Comparison to simulations with site-wise dominance.}
        \(Ne=5000\), \(r=5\times10^{-6}\), and \(s=-0.0002\) (\(\rho=0.1\) and
        \(\gamma=-2\)), with \(h=0.1\) at both loci and no
        epistatic interactions (\(\epsilon=0\)).
        E: \texttt{moments} expectations, D: simulated data.
        Additional comparisons to simulated data are available at
        \texttt{https://github.com/apragsdale/two\_locus\_selection}.
    }
    \label{fig:validation5}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with slightly deleterious mutations.}
        Using an individual-based forward simulator \citep{Thornton2019-qc}, I
        simulated many replicates of 1Mb regions with a high mutation rate of
        selected mutations (see Section~\ref{sec:bgs} for details). Here,
        many linked additive weakly selected mutations slightly distort the
        SFS for both (A) selected and (C) neutral mutations.
        (B) In this scenario of background selection, \(\sigma_d^2\)
        is reduced relative to two-locus expectations without linked selection.
        (D) Hill-Robertson interference effects are slightly diminished, with
        \(\sigma_d^1\) not as strongly negative as expected without additional
        linked selected mutations.
    }
    \label{fig:bgs1}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.5_n_50}
    \caption{
        \textbf{Background selection with moderately deleterious mutations.}
        In the case of many linked strongly selected additive mutations, neither
        selected nor neutral statistics deviate dramatically from expectations
        for either the SFS (A, C) or two-locus statistics (B, D).
        The SFS from the simulated data matches single-locus expectations.
        LD is also largely unaffected, with \(\sigma_d^2\)
        matching expectations and \(\sigma_d^1\) fluctuating near zero as expected.
    }
    \label{fig:bgs2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-1.0_h_0.0_n_50}
    \caption{
        \textbf{Associative overdominance due to slightly deleterious recessive mutations.}
        The presence of many linked weakly selected recessive mutations can cause
        measures of diversity to have large increases over single- and two-locus
        expectations. The SFS for both selected (A) and neutral (C) mutations both show
        a very large excess of common variants (note the log scale), and \(\sigma_d^2\)
        decay curves (B)
        are also increased over two-locus expectations for both classes of mutations.
        (D) \(\sigma_d^1\) remains 0 for neutral mutations and is not as strongly
        negative for pairs of selected mutations due to interference from linked selected
        mutations.
    }
    \label{fig:bgs3}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/bgs_gamma_-20.0_h_0.1_n_80}
    \caption{
        \textbf{Background selection with moderately deleterious recessive mutations.}
        More strongly deleterious recessive mutations do not show as dramatic of
        deviations from expectations as the case of weakly deleterious recessive
        mutations (Fig.~\ref{fig:bgs3}.
        (A) We observe more selected mutations at higher frequencies than expected,
        and (C) the neutral SFS is reduced below single-locus expectations.
        (B, D) LD is also only weakly distorted from
        two-locus expectations. Notably, we see positive LD for strongly
        deleterious recessive mutations (which matches between
        simulations and expectations),
        as previously predicted in \citet{Roze2021-cf}.
    }
    \label{fig:bgs4}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.sd2}
    \caption{
        \textbf{Time series of squared LD under bottleneck 
        and expansion histories with epistasis.}
        Bottlenecks increase \(\sigma_d^2\) (squared LD),
        while expansions tend to decrease squared LD.
        Black dashed lines show neutral expectations.
        (A) The simulated size histories. (B-F) Blue solid lines
        are expected squared LD under the expansion model, and gold
        dashed lines are squared LD under the bottleneck/recovery model.
        (B-D) Expectations are computed over all allele frequencies.
        (E) Expected squared LD between uncommon variants, with $n_A$
        and $n_B \leq 4$, in a sample size of 50 haploid copies.
        (F) Expected LD squared LD between common variants
        $p_A, p_B \gtrsim 10\%$.
    }
    \label{fig:toy_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance}
    \caption{
        \textbf{Time series of signed LD under bottleneck 
        and expansion histories with dominance.}
        Gene-based dominance, unlike site-wise dominance effects, can cause
        large positive signed LD, similar to compensatory mutations or
        antagonistic epistasis.
        (A) The simulated size histories.
        (B-D) Signed LD computed from pairs of mutations
        across all allele frequencies.
        (E, F) In general, common variants are more
        stable under population size changes than rare variants.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_dom}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_bottle_expand.dominance.sd2}
    \caption{
        \textbf{Time series of squared LD under bottleneck 
        and expansion histories with dominance.}
        Similar to the cases with epistasis, bottlenecks tend to increase
        \(\sigma_d^2\) and expansions decrease squared LD.
        (A) The simulated size histories.
        (B-D) Squared LD computed from pairs of mutations at all allele frequencies.
        (E, F) Squared LD computed from uncommon ($<10\%$) and common ($\gtrsim 10\%$)
        mutations.
        Dashed lines show neutral expectations.
    }
    \label{fig:toy_dom_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU}
    \caption{
        \textbf{Time series of signed LD under inferred models of human population
        size history with with epistasis.}
        (A) Size histories inferred using Relate \citep{Speidel2019-nj}.
        (B-F)
        Similar to the instantaneous size changes in the toy models
        (Figure~\ref{fig:toy}), bottlenecks tend to make signed LD more extreme.
        However, in the case of strong selection and antagonistic epistasis (D),
        the CEU \(\sigma_d^1\) trajectory decreases relative to the YRI
        trajectory, showing that non-constant demographic history and epistasis can
        have complicated interactions.
        (E, F) Again, statistics for rare variants fluctuate more rapidly
        than common variants.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.sd2}
    \caption{
        \textbf{Time series of squared LD under inferred models of human population
        size history with with epistasis.}
        (A) Inferred size histories used in simulations.
        (B-D)
        Size reductions, such as bottlenecks in the recent past of Eurasian populations,
        increase squared LD.
        (E, F) Uncommon and common variants can show strikingly different behaviors
        under size change histories, here showing trajectories under the three
        epistasis models from panels B-D.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate_sd2}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/demog_YRI_CEU.dominance.sd2}
    \caption{
        \textbf{Time series of squared LD under inferred models of human population
        size history with with dominance.}
        (A) Inferred size histories used in simulations.
        (B-D)
        Again, bottlenecks increase squared LD relative to populations that did
        not experience such strong size reductions in their history.
        (E, F) Uncommon and common variants can show strikingly different behaviors
        under size change histories, here showing trajectories under the three
        dominance models from panels B-D.
        Dashed lines show neutral expectations.
    }
    \label{fig:relate_dom_sd2}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/ld_decay_in_domain}
    \caption{
        \textbf{LD decay for pairs of synonymous and missense mutations
        that fall inside the same domain.}
        At short distances between SNPs that both fall within sthe same conserved
        annotated domain, missense mutations show consistently
        larger signed LD than synonymous variants. This may be caused by
        antagonistic epistasis or compensatory effects between pairs of
        nearby missense mutations.
    }
    \label{fig:LDwithin}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_afr}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes African group.
        We observe increased LD between missense mutations within domains,
        but decreased LD outside of domains at matched distances. This pattern
        holds when conditioning on non-rare variation (\(n_A, n_B \geq 9\),
        MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsAFR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eur}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes European group.
        We again observe increased LD between missense mutations within domains,
        but decreased LD outside of domains at matched distances. And again,
        this pattern holds when conditioning on non-rare variation
        (\(n_A, n_B \geq 9\), MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsEUR}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics{../figures/data_domains_eas}
    \caption{
        \textbf{LD for pairs of synonymous and missense mutations within the
        same domain and outside domains at matched distances.}
        Showing the five populations in the Thousand Genomes East Asian group.
        As with the African and European data,
        we observe increased LD between missense mutations within domains,
        but decreased LD outside of domains at matched distances. And again,
        this pattern holds when conditioning on non-rare variation
        (\(n_A, n_B \geq 9\), MAF \(\gtrsim 0.05\)).
    }
    \label{fig:domainsEAS}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{../figures/SFS_proportions}
    \caption{
        \textbf{SFS for common missense and synonymous variants.}
        The full SFS (considering variants at all observed allele frequencies) is
        skewed toward low frequencies for missense variants compared to synonymous
        variants (shown in light colors). Tajima's $D$ is considerably more negative
        for missense variants than synonymous variants for missense variants, whether
        conditioning on mutations within annotated elements or outside
        (Table~\ref{tab:tajimasD}). However, the SFS for common variants (with
        observed MAF $\geq 10\%$) are similar between missense and synonymous variants
        for each partition of the data. Thus, the excess of positive signed LD for
        common variants within annotated domains is likely not to be driven by
        subtle differences in allele frequencies between missense and synonymous
        classes.
    }
    \label{fig:sfs_proportions}
\end{figure}
